---
title: "Covid19_Uk"
author: "FG"
date: "8/9/2020"
output: html_document
---
```{r setup, include=FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(tidyverse)
library(devtools)
library(deSolve)#to solve differential equations
library(zoo)#for rolling averages
library(slider)#for rolling averages
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
dataECDC <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", na.strings = "", fileEncoding = "UTF-8-BOM")

```

```{r}
names(dataECDC)
```


Table of the last 6 days of COVID19 UK data:

```{r, echo=FALSE}

Covid19_UK<- dataECDC%>%
  filter(countriesAndTerritories=="United_Kingdom")%>%
  dplyr::select(1,10,5,6,12)%>%
  mutate(Date=dateRep,Cum_14days=round(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000,3))%>%
  dplyr::select(Date,popData2019,cases,deaths,Cum_14days)%>%
  mutate(Days=c(length(Date):1),survivors=cases-deaths,incidence=round((cases-deaths)/cases,2),CMR_100=round((deaths/popData2019)*100*100000,2),CFR_100=round((deaths/cases)*100,2))%>%
  dplyr::select(Date,Days,cases,survivors,deaths,Cum_14days,incidence,CFR_100,CMR_100)
 
head(Covid19_UK,6)
```
```{r}
require(zoo)
ggplot(Covid19_UK, aes(Days, cases)) + 
  geom_point(position=position_jitter(1,3), pch=21, fill="#FF0000AA") +
  geom_line(aes(y=rollmean(cases, 7, na.pad=TRUE))) +
  theme_bw()
```


```{r}


ggplot(Covid19_UK, aes(Days, incidence)) + 
  geom_point(position=position_jitter(1,3), pch=21, fill="#FF0000AA") +
  geom_line(aes(y=rollmean(incidence, 7, na.pad=TRUE))) +
  theme_bw()

```

Data are sparse and is necessary a further analysis to see if the behavior is 
Normal. To do that:

```{r}
       
Covid19_UK$incidence[is.nan(Covid19_UK$incidence)]<- 0   #elimination of NaN values
incidence<- Covid19_UK$incidence #creating a incidence variable

n<- length(incidence)
inc_avg<- mean(incidence)
inc_sd<- sd(incidence)
inc_se<- inc_sd/sqrt(n) 

inc_norm<- pnorm((incidence-inc_avg)/inc_sd)
plot(density(inc_norm))


```

```{r}
pnorm(15,0,1)#changes
```


This is a second way to calculate a rolling average:

```{r}
library(slider)

prevision<- Covid19_UK %>% 
  mutate(rolling_avg = slide_dbl(cases, mean, .before = 3, .complete = F))


  ggplot()+
  geom_line(data=Covid19_UK,aes(x=Days,y=cases))+
  geom_line(data=prevision,aes(x=Days,y=rolling_avg),col="yellow")
```


```{r}

ggplot()+
  geom_histogram(data=Covid19_UK, aes(x=cases),bins=30, fill = "darkgreen", 
                 col = "black")+
  labs(x="Cases", y="Frequency",title="UK Cases")+
  scale_x_continuous(breaks = seq(from = 0, to = 7000, by = 500)) + theme_bw()

```

Table of the last 6 days of COVID19 Italy data:


```{r, echo=FALSE}
Covid19_IT<- dataECDC%>%
  filter(countriesAndTerritories=="Italy")%>%
  dplyr::select(dateRep,popData2019,cases,deaths,Cumulative_number_for_14_days_of_COVID.19_cases_per_100000)%>%
  mutate(Date=dateRep,Cum_14days=round(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000,0))%>%
  dplyr::select(Date,popData2019,cases,deaths,Cum_14days)%>%
  mutate(Days=c(length(Date):1),survivors=cases-deaths,incidence=round((cases-deaths)/cases,2),CMR_100=round((deaths/popData2019)*100*100000,2),CFR_100=round((deaths/cases)*100,2))%>%
  dplyr::select(Date,Days,cases,survivors,deaths,Cum_14days,incidence,CFR_100,CMR_100)
 
head(Covid19_IT,6)
```
Plot of the two curves of cumulated cases for UK and Italy for the whole period:

```{r}
ggplot()+
  geom_line(data=Covid19_UK,aes(x=Days,y=cases),col="blue",size = 0.3)+
  geom_line(data=Covid19_IT,aes(x=Days,y=cases),col="red",size = 0.3)+
  annotate(geom="text", x=50, y=6000, label="Italy",
              color="red")+
  annotate(geom="text", x=150, y=4000, label="UK",
              color="blue")+
    theme_bw(base_size = 12) + theme(legend.position = "bottom")+
  xlab("Time(days)")+ylab("Cases")+
  labs("UK and Italy cumulated cases trend",color = "Legend")
  
```

UK Cases, Deaths and Survivors trends: 

```{r}

ggplot()+
  geom_line(data=Covid19_UK,aes(x=Days,y=cases),col="green",size=0.3)+
  geom_line(data=Covid19_UK,aes(x=Days,y=deaths),col="red",size=0.3)+
  geom_line(data=Covid19_UK,aes(x=Days,y=survivors),col="blue",size=0.3)+
  theme_bw(base_size = 12) + theme(legend.position = "bottom")+
  labs(x="Time(days)", y="Cumulate numbers", title="UK Cases, deaths and CFR trends")
 
 
```


UK Crude Mortality rates (CMR) and Case Fatality Rates (CFR):

```{r}

plot(density(Covid19_UK$CMR_100),col="green",xlim=c(-20,50),ylim=c(0,0.1),main="UK CMR and CFR")
lines(density(Covid19_UK$CFR_100,na.rm = T),col="red")
axis(side = 1, at = round(seq(0, 50, by = 10)))
text(x=median(Covid19_UK$CMR_100), y=0.04, labels="CMR")
text(x=mean(Covid19_UK$CFR_100,na.rm=T), y=0.07, labels="CFR")
```


The CFR curve shows a right(positive) skewed distribution of fatalities along the time:


```{r}
hist(Covid19_UK$CFR_100,main="UK CFR")
```


In this section is represented the basic of the SIR model with given beta and gamma,
it is expected a further calibration of the model considering other factors, such as:
- effect of treatment
- effect of social distancing
- age impact
- co-morbidity
- symptomatic and asymptomatic behavior
- vaccination strategies

To begin with a theoretical value of beta to calibrate the model using a simple linear
regression model function: the sum of squares values (SSQ). The line that best predict the lowest 
distance between theoretical predictors and real data, to assess how far is the model from real data.


The prevalence of the theoretical model infection is represented in the graph below:

```{r}
N<- max(Covid19_UK$cases)

initial_state_values<- c(S=N-1,I=1,R=0)

parameters<- c(beta= 1/3,gamma=1/22)#beta (infection rate) is hypothesized to be 3 days^-1, 
                                    #gamma (recovery rate) is 22 days^-1

times<- seq(0,length(Covid19_UK$Date),by=1)

SIR_model<- function(time,state,parameters){
  with(as.list(c(state,parameters)),{
       N<- S+I+R
       lambda<- beta*I/N
       dS<- -lambda*S
       dI<- lambda*S-gamma*I
       dR<- gamma*I
       return(list(c(dS,dI,dR)))
       })
}

output<- as.data.frame(ode(y=initial_state_values,
                           times=times,
                           func=SIR_model,
                           parms=parameters))

ggplot()+
  geom_line(data=output,aes(x=time,y=I))+
  labs(x="Time(days)",y="Infected", title="UK Prevalence of infection")

```





```{r}
output_long<- melt(as.data.frame(output), id="time")
ggplot(data=output_long, aes(x=time, y=value,colour=variable,group=variable))+
  geom_line()+
 labs( x="Time(days)",y="proportions", title="UK Covid19 theoretical Prevalence proportion: basic SIR model")
  
```

Simple Linear Regression:

```{r}

SIR_fun<- function(time,state,parameters){
  with(as.list(c(state,parameters)),{
       N<- S+I+R
      
       dS<- -beta*S*(I/N)
       dI<- beta*S*(I/N)
       dR<- gamma*I
       return(list(c(dS,dI,dR)))
       })}
  
SIR_SSQ <- function(parameters, dat) {  
            
    result <- as.data.frame(ode(y = initial_state_values  
                              , times = times            
                              , func = SIR_fun           
                              , parms = parameters)      
                            )
    dat <- na.omit(dat)  
    
    deltas2 <- (result$I[result$time %in% dat$time] - dat$I)^2 
    
    SSQ   <- sum(deltas2)
    
    return(SSQ)
  }

```




```{r}
require(reshape2)

dat_UK<- Covid19_UK%>%
  mutate(number_infected = (cases-deaths),time=Days)%>%
  dplyr::select(time,number_infected)

head(dat_UK)
```


```{r}
initial_state_values <- c(S = 10500-1, I = 1, R = 0)

beta_start  <- 0.15
gamma_start <- 0.04

times <- seq(from = 0, to = 14, by = 0.1) 

optimised <- optim(par = c(beta = beta_start
                        , gamma = gamma_start)     
                        , fn = SIR_SSQ
                        , dat = dat_UK  
      )

```

```{r}
round(optimised$par,2)
```


```{r}
times <- seq(from = 0, to = length(Covid19_UK$cases), by = 1) 

opt_mod <- as.data.frame(ode(y = initial_state_values,
                             times = times,
                             func = SIR_model,
                             parms = optimised$par))

```

```{r}
require(ggplot2)


opt_plot <- ggplot()
opt_plot <- opt_plot + geom_point(aes(x = time, y = number_infected)
                                , colour = "red"
                                , shape  = "x" 
                                , data = dat_UK)
opt_plot <- opt_plot + geom_line(aes(x = time, y = I)
                                 , colour = "blue"
                                 , data   = opt_mod)
opt_plot
```







